---
title: "NewResults"
format: html
editor: visual
---

#Drivers of EFN plant richness

```{r}
library(sf)
library(raster)
library(dplyr)
library(tidyr)
library(ggplot2)
```

```{r}
#stack the prediction rasters

efnlist<- raster::stack("data/output/predictions/mean_pred_CA.grd", "data/output/predictions/mean_pred_CE.grd", "data/output/predictions/mean_pred_AG.grd", "data/output/predictions/mean_pred_FS.grd", "data/output/predictions/mean_pred_CL.grd", "data/output/predictions/mean_pred_OB.grd", "data/output/predictions/mean_pred_OBIG.grd", "data/output/predictions/mean_pred_FA.grd", "data/output/predictions/mean_pred_PJ.grd", "data/output/predictions/mean_pred_PFA.grd", "data/output/predictions/mean_pred_PFREM.grd")

names(efnlist) <- c("CA", "CE", "AG", "FS", "CL", "OB", "OBIG", "FA", "PJ", "PFA", "PFREM")

plot(efnlist)
EFN_sum <- sum(efnlist, na.rm = TRUE)
plot(EFN_sum)


#clip to mojave
r <- brick("na_rasters.grd")

crs <- st_crs(r)
moj <- read_sf("data/mojavescol.shp")
moj <- st_transform(moj, crs)

EFN_sum <- mask(EFN_sum, moj)
efnlist <- mask(efnlist, moj)
plot(EFN_sum)
plot(efnlist)

clim <- stack(
      "data/output/predictions/mean_pred_CAMPFRA.grd",
      "data/output/predictions/mean_pred_CAMPOC.grd",
      "data/output/predictions/mean_pred_CAMPSEM.grd",
      "data/output/predictions/mean_pred_MYRMIM.grd",
      "data/output/predictions/mean_pred_CREMD.grd",
      "data/output/predictions/mean_pred_DORBI.grd", 
      "data/output/predictions/mean_pred_DORIN.grd",
      "data/output/predictions/mean_pred_MYRFLA.grd",
      "data/output/predictions/mean_pred_MYRKEN.grd",
      "data/output/predictions/mean_pred_MYRTES.grd",      
      "data/output/predictions/mean_pred_SX.grd",
      "data/output/predictions/mean_pred_LIOLUC.grd",
      "data/output/predictions/mean_pred_PHEVIS.grd",      
      "data/output/predictions/mean_pred_FM.grd")

names <- c(
      "CAMPFRA.grd",
      "CAMPOC.grd",
      "CAMPSEM.grd",
      "MYRMIM.grd",
      "CREMD.grd",
      "DORBI.grd", 
      "DORIN.grd",
      "MYRFLA.grd",
      "MYRKEN.grd",
      "MYRTES.grd",      
      "SX.grd",
      "LIOLUC.grd",
      "PHEVIS.grd",
      "FM.grd")

names(clim) <- names

ants <- sum(clim)
ants <- mask(ants, moj)
clim <- mask(clim, moj)
plot(ants)
plot(clim)


```

## Evaluate models

```{r}


#ants 
filenames <- list.files("predictions/ants", full.names = TRUE, pattern = "csv$")


ants <- purrr::map_df(filenames, 
              ~read.csv(.x, stringsAsFactors = FALSE) %>% mutate(filename = .x))


ants_mods <- ants %>% group_by(filename) %>% summarise(auc = mean(test_auc), aucsd = sd(test_auc), tss = mean(test_tss), tsssd = sd(test_tss))

ants_mods[,2:5] <- round(ants_mods[,2:5], 2)




filenames <- list.files("predictions/antcontrasts", full.names = TRUE, pattern = "csv$")


ants_contrasts <- purrr::map_df(filenames, 
              ~read.csv(.x, stringsAsFactors = FALSE) %>% mutate(filename = .x))



#plants 
filenames <- list.files("predictions/", full.names = TRUE, pattern = "csv$")


plants <- purrr::map_df(filenames, 
              ~read.csv(.x, stringsAsFactors = FALSE) %>% mutate(filename = .x))


plants_mods <- plants %>% group_by(filename) %>% summarise(auc = mean(test_auc), aucsd = sd(test_auc), tss = mean(test_tss), tsssd = sd(test_tss))

plants_mods[,2:5] <- round(plants_mods[,2:5], 3)


write.csv(plants_mods, "plantCV.csv")


```

## Collect EFN models

```{r}
#ants is clim only models

ants$type <- "clim"
ants <- rename(ants, species = filename)
ants$species <- gsub("predictions/ants/CV_Results_", "", ants$species) 
ants$species <- gsub(".csv", "", ants$species) 


#ants_contrasts
ants_contrasts$filename <- gsub("predictions/antcontrasts/CV_Results_", "", ants_contrasts$filename) 
ants_contrasts$filename <- gsub(".csv", "", ants_contrasts$filename) 
ants_contrasts <- separate(ants_contrasts, filename, into = c("species", "type", "bin"), sep = "_")
ants_contrasts$bin[is.na(ants_contrasts$bin)] <- "con"

ants_contrasts <- filter(ants_contrasts, bin != "bin")
ants_contrasts <- filter(ants_contrasts, type == "EFN" | type == "EFN-only" | type == "EFN_only")




ants_contrasts <- mutate(ants_contrasts, treat = ifelse(bin == "only", "EFN-only", type))


ants_contrasts <- select(ants_contrasts, -bin, - type)

ants_contrasts <- rename(ants_contrasts, type = treat)

dfcomp <- rbind(ants, ants_contrasts)
dfcomp <- filter(dfcomp, species != "CAMPSAY" & species != "FP")
```

## Compare model performance

```{r}

library(glmmTMB)
library(ggplot2)
library(performance)
dfcomp2 <- filter(dfcomp, type != "EFN-only")

m1 <- glmmTMB(test_auc ~ type + (1|species), data = dfcomp)
model_performance(m1)
summary(m1)
plot(resid(m1))

ggplot(dfcomp2, aes(type, test_auc)) + geom_boxplot()

ggplot(dfcomp, aes(species, test_auc)) + geom_boxplot(aes(fill = type)) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black")) + ylab("Test AUC (predictive accuracy)") + xlab("Ant Species") + scale_fill_manual(values = c("gray", "black", "white"), labels = c("Climate Only", "EFN + Climate", "EFN Only")) + stat_summary(fun.y=mean, aes(group = type, fill = type), geom="point", size=1, color = "red", position=position_dodge(width=0.75)) +
scale_x_discrete(labels = c("Camponotus \nfragilis", "Camponotus \nocreatus", "Camponotus \nsemitestaceus", "Crematogaster \ndepilis", "Dorymyrmex \nbicolor", "Dorymyrmex \ninsanus", "Forelius \nmccooki", "Liometopum \nluctuosum", "Myrmecocystus \nflaviceps", "Myrmecocystus \nkennedyi", "Myrmecocystus \nmimicus", "Myrmecocystus \ntestaceus", "Pheidole \nvistana", "Solenopsis \nxyloni")) +
  theme(legend.title=element_blank()) + theme(axis.text.x = element_text(angle = 90, hjust = 1))


dfcomp %>% group_by(species, type) %>% summarize(mean = mean(test_auc))
```

plot(efnlist\[\[1\]\]) plot(ca_thin, add = TRUE)

plot(efnlist)

## Environmental Rasters

```{r}

r <- brick("na_rasters.grd")
# #
# #
crs <- st_crs(r)
moj <- read_sf("data/mojavescol.shp")
moj <- st_transform(moj, crs)
e = terra::ext(moj)

env.nopca <- crop(r, extent(moj))


#check occurrences using mcp
env.nopca <- mask(env.nopca, moj)
env.nopca <- terra::rast(env.nopca)


```

## Plant Future

```{r}

efn245<- raster::stack("data/output/predictions/Future/ssp245/mean_pred_CA_ssp245.grd", "data/output/predictions/Future/ssp245/mean_pred_CE_ssp245.grd", "data/output/predictions/Future/ssp245/mean_pred_AG_ssp245.grd", "data/output/predictions/Future/ssp245/mean_pred_FS_ssp245.grd", "data/output/predictions/Future/ssp245/mean_pred_CL_ssp245.grd", "data/output/predictions/Future/ssp245/mean_pred_OB_ssp245.grd", "data/output/predictions/Future/ssp245/mean_pred_OBIG_ssp245.grd", "data/output/predictions/Future/ssp245/mean_pred_FA_ssp245.grd", "data/output/predictions/Future/ssp245/mean_pred_PJ_ssp245.grd", "data/output/predictions/Future/ssp245/mean_pred_PFA_ssp245.grd", "data/output/predictions/Future/ssp245/mean_pred_PFREM_ssp245.grd")

plot(efn245)
efn245_sum <- sum(efn245)

efn370<- raster::stack("data/output/predictions/Future/ssp370/mean_pred_CA_ssp370.grd", "data/output/predictions/Future/ssp370/mean_pred_CE_ssp370.grd", "data/output/predictions/Future/ssp370/mean_pred_AG_ssp370.grd", "data/output/predictions/Future/ssp370/mean_pred_FS_ssp370.grd", "data/output/predictions/Future/ssp370/mean_pred_CL_ssp370.grd", "data/output/predictions/Future/ssp370/mean_pred_OB_ssp370.grd", "data/output/predictions/Future/ssp370/mean_pred_OBIG_ssp370.grd", "data/output/predictions/Future/ssp370/mean_pred_FA_ssp370.grd", "data/output/predictions/Future/ssp370/mean_pred_PJ_ssp370.grd", "data/output/predictions/Future/ssp370/mean_pred_PFA_ssp370.grd", "data/output/predictions/Future/ssp370/mean_pred_PFREM_ssp370.grd")

plot(efn370)
efn370_sum <- sum(efn370)


```

## Ant Future

```{r}

ant245 <- stack(
      "data/output/predictions/Future/ssp245/mean_pred_CAMPFRA_ssp245.grd",
      "data/output/predictions/Future/ssp245/mean_pred_CAMPOC_ssp245.grd",
      "data/output/predictions/Future/ssp245/mean_pred_CAMPSEM_ssp245.grd",
      "data/output/predictions/Future/ssp245/mean_pred_MYRMIM_ssp245.grd",
      "data/output/predictions/Future/ssp245/mean_pred_CREMD_ssp245.grd",
      "data/output/predictions/Future/ssp245/mean_pred_DORBI_ssp245.grd", 
      "data/output/predictions/Future/ssp245/mean_pred_DORIN_ssp245.grd",
      "data/output/predictions/Future/ssp245/mean_pred_MYRFLA_ssp245.grd",
      "data/output/predictions/Future/ssp245/mean_pred_MYRKEN_ssp245.grd",
      "data/output/predictions/Future/ssp245/mean_pred_MYRTES_ssp245.grd",      
      "data/output/predictions/Future/ssp245/mean_pred_SX_ssp245.grd",
      "data/output/predictions/Future/ssp245/mean_pred_LIOLUC_ssp245.grd",
      "data/output/predictions/Future/ssp245/mean_pred_PHEVIS_ssp245.grd",
      "data/output/predictions/Future/ssp245/mean_pred_FM_ssp245.grd")


plot(ant245)






ant245_sum <- sum(ant245)

ant370 <- stack(
      "data/output/predictions/Future/ssp370/mean_pred_CAMPFRA_ssp370.grd",
      "data/output/predictions/Future/ssp370/mean_pred_CAMPOC_ssp370.grd",
      "data/output/predictions/Future/ssp370/mean_pred_CAMPSEM_ssp370.grd",
      "data/output/predictions/Future/ssp370/mean_pred_MYRMIM_ssp370.grd",
      "data/output/predictions/Future/ssp370/mean_pred_CREMD_ssp370.grd",
      "data/output/predictions/Future/ssp370/mean_pred_DORBI_ssp370.grd", 
      "data/output/predictions/Future/ssp370/mean_pred_DORIN_ssp370.grd",
      "data/output/predictions/Future/ssp370/mean_pred_MYRFLA_ssp370.grd",
      "data/output/predictions/Future/ssp370/mean_pred_MYRKEN_ssp370.grd",
      "data/output/predictions/Future/ssp370/mean_pred_MYRTES_ssp370.grd",      
      "data/output/predictions/Future/ssp370/mean_pred_SX_ssp370.grd",
      "data/output/predictions/Future/ssp370/mean_pred_LIOLUC_ssp370.grd",
      "data/output/predictions/Future/ssp370/mean_pred_PHEVIS_ssp370.grd",
      "data/output/predictions/Future/ssp370/mean_pred_FM_ssp370.grd")
plot(ant370)
ant370_sum <- sum(ant370)
plot(ant370)
plot(ant370_sum)
```

## Env Future

```{r}
ssp245 <- brick("ssp245.grd")

ssp245 <- terra::rast(ssp245)
ssp370 <- brick("ssp370.grd")
ssp370 <- terra::rast(ssp370)
plot(ssp245)
plot(ssp370)


```

## Environmental Shifts

PC1 and PC2 Shifts

```{r}

envs <- brick("envs.grd")
plot(envs)
env <- mask(envs, moj)
plot(env)
pc1pres <- env[[1]]
plot(pc1pres)
ssp245 <- mask(ssp245, moj)
ssp370 <- mask(ssp370, moj)
pc1245 <- ssp245[[1]]
pc1pres <- terra::rast(pc1pres)
plot(pc1pres)
plot(pc1245)

pc1change <- pc1pres - pc1245
plot(pc1change)

pc2pres <- env[[2]]
pc2pres <- terra::rast(pc2pres)
pc2245 <- ssp245[[2]]
pc2change <- pc2pres - pc2245
plot(pc2change)
plot(pc2pres)


shift <- terra::rast(env) - ssp245
plot(shift)

shift2<- ssp370
plot(shift2)

plot(ssp370[[4]])
plot(env)
```

## Richness sampling

```{r}
ants <- sum(clim)
ants <- mask(ants, moj)
plot(ants)

env <- mask(envs, moj)
plot(env)
all <- stack(EFN_sum, ants, env)
names(all)
names(all) <- c("EFN", "ants", "PC1", "PC2", "PC3", "PC4")



plot(all)
#par(mfrow = c(1, 2))
plot(ants)
plot(EFN_sum)



# 
# val <- values(all)
# val <- as.data.frame(all)
# val <- drop_na(val)

pts <- dismo::randomPoints(all, 5000)
df <- raster::extract(all, pts)
df <- as.data.frame(df)

dfenv <- df[,3:6]



library(vegan)
v1 <- varpart(df$ants, df$EFN, dfenv)
plot(v1)


```

## Change Analyses

```{r}


EFN_sum <- mask(EFN_sum, moj)
plot(EFN_sum)
efn245_sum <- mask(efn245_sum, moj)
plot(efn245_sum)
efn370_sum <- mask(efn370_sum, moj)
plot(efn370_sum)

efnchange <- stack(EFN_sum, efn245_sum, efn370_sum)
names(efnchange) <- c("EFN", "245", "370")


# 
# val <- values(all)
# val <- as.data.frame(all)
# val <- drop_na(val)

pts <- dismo::randomPoints(efnchange, 5000)
df <- raster::extract(efnchange, pts)
df <- as.data.frame(df)
df$row <- row.names(df)

dfefn <- pivot_longer(df, 1:3, names_to = "scenario", values_to = "rich")

library(glmmTMB)
m1 <- glmmTMB(rich ~ scenario + (1|row), data = dfefn)
summary(m1)
car::Anova(m1)
library(emmeans)

emmeans(m1, pairwise ~ scenario)

ggplot(dfefn, aes())

plot(ants)
ant245_sum <- mask(ant245_sum, moj)
plot(ant245_sum)
ant370_sum <- mask(ant370_sum, moj)
plot(ant370_sum)

antchange <- stack(ants, ant245_sum, ant370_sum)
names(antchange) <- c("ant", "245", "370")

plot(antchange)

# 
# val <- values(all)
# val <- as.data.frame(all)
# val <- drop_na(val)

pts <- dismo::randomPoints(antchange, 5000)
df <- raster::extract(antchange, pts)
df <- as.data.frame(df)
df$row <- row.names(df)

dfant <- pivot_longer(df, 1:3, names_to = "scenario", values_to = "rich")

library(glmmTMB)
m1 <- glmmTMB(rich ~ scenario + (1|row), data = dfant)
summary(m1)
emmeans(m1, pairwise ~ scenario)

library(ggplot2)



dfefn$type <- "EFN"
dfant$type <- "ants"

all <- rbind(dfefn, dfant)
ggplot(all, aes(type, rich, fill = scenario)) + geom_boxplot()

m1 <- glmmTMB(rich ~ scenario*type + (1|row), data = all)


#maps for losses
antloss <- ant370_sum - ants
par(mfrow=c(1,1))
plot(antloss)

efnloss <- efn370_sum - EFN_sum
pal <- colorRampPalette(c("white","black"))

df = as.data.frame(efnloss, xy=TRUE)

dfantloss = as.data.frame(antloss, xy=TRUE)

par(mfrow=c(2,2))
#, values = c(-4.381, 0, 3.2908)
ggplot(df) + geom_raster(aes(y=y, x=x, fill=layer)) + scale_fill_gradient2(low = "red",high = "darkgreen") +  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black")) + xlab("Longitude") + ylab("Latitude") + labs(fill = "Absolute change in richness")

ggplot(dfantloss) + geom_raster(aes(y=y, x=x, fill=layer)) + scale_fill_gradient2(low = "red",high = "darkgreen") +  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black")) + xlab("Longitude") + ylab("Latitude") + labs(fill = "Absolute change in richness")
par(mfrow=c(1,1))
```


## Suitability shifts

```{r}

ant245 <- mask(ant245, moj)
efn245 <- mask(efn245, moj)
ant370 <- mask(ant370, moj)
efn370 <- mask(efn370, moj)
plot(clim)
plot(efnlist)
plot(efn370)

diff_ants_245 <- ant245 - clim
diff_efn_245 <- efn245 - efnlist

plot(diff_ants_245)
plot(diff_efn_245)

clim
cellStats(diff_ants_245, "mean")

efnlist
cellStats(diff_efn_245, "mean")


names(efnlist)
names(efn245)

diff_ants_370 <- ant370 - clim
diff_efn_370 <- efn370 - efnlist

plot(diff_ants_370)
plot(diff_efn_370)

clim
cellStats(diff_ants_370, "mean")
cellStats(diff_efn_370, "mean")


ant_diff <- rbind(names(clim), cellStats(diff_ants_245, "mean"),  cellStats(diff_ants_370, "mean")) 

row.names(ant_diff) <- c("names", "diff245", "diff370")


efn_diff <- rbind(names(efnlist), cellStats(diff_efn_245, "mean"), cellStats(diff_efn_370, "mean")) 

row.names(efn_diff) <- c("names", "diff245", "diff370")



```

## Suitability shift map

```{r}

shift370 <- ant370_sum - ants
plot(shift370)
df = as.data.frame(shift370, xy=TRUE)
ggplot(df) + geom_raster(aes(y=y, x=x, fill=layer)) + scale_fill_gradient2(low = "red",high = "darkgreen", name="Change in \nSpecies Richness") +  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"))


efn_change_370 <- efn370_sum - EFN_sum
df = as.data.frame(efn_change_370, xy=TRUE)
ggplot(df) + geom_raster(aes(y=y, x=x, fill=layer)) + scale_fill_gradient2(low = "red",high = "darkgreen", name="Change in \nSpecies Richness") + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black")) 

plot(EFN_sum)
plot(ants)
plot(efn370_sum)
plot(ant370_sum)





```

# Phylogeny

```{r}

#install.packages("phytools")
#install.packages("picante")
library(picante)
library(phytools)
library(tidyr)
genusv <- c("Camponotus", "Crematogaster", "Dorymyrmex", "Forelius", "Liometopum", "Myrmecocystus", "Pheidole", "Solenopsis")

#import nexus tree
tree <- read.nexus("data/moreaubeast.nex")
str(tree)
plot(tree)
tree
labels <- tree[["tip.label"]]
labels <- as.data.frame(labels)
full <- labels$labels
labels <- separate_wider_delim(labels, cols = 1, delim ="_", names = c("genus", "sp", "csm"), too_few = "align_start", too_many = "merge")
labels <- cbind(full, labels)
labels <- subset(labels, genus %in% genusv)
keep <- labels$full
df <- data.frame(matrix(ncol = 21, nrow = 1))
colnames(df) <- keep
df[is.na(df)] <- 1
df <- dplyr::select(df, 1, 4, 7, 10, 13, 16, 18, 21)
pruned <- prune.sample(df, tree)
pruned <- add.species.to.genus(pruned, "Myrmecocystus_kennedyi")
pruned <- add.species.to.genus(pruned, "Myrmecocystus_testaceus")
pruned <- add.species.to.genus(pruned, "Myrmecocystus_mimicus")
pruned <- add.species.to.genus(pruned, "Camponotus_fragilis")
pruned <- add.species.to.genus(pruned, "Camponotus_semitestaceus")
pruned <- add.species.to.genus(pruned, "Dorymyrmex_insanus")

pruned
pruned[["tip.label"]]
pruned[["tip.label"]] <- c("CREMD", "PHEVIS", "SX", "MYRMIM", "MYRTES", "MYRFLA", "MYRKEN", "CAMPSEM", "CAMPOC", "CAMPFRA", "DORBI", "DORIN", "FM", "LIOLUC") 
plot(pruned)

labelorder <- c("CREMD", "PHEVIS", "SX", "MYRMIM", "MYRTES", "MYRFLA", "MYRKEN", "CAMPSEM", "CAMPOC", "CAMPFRA", "DORBI", "DORIN", "FM", "LIOLUC")
plot(pruned)
is.ultrametric(pruned)

```


## Schoener's D Calculations

```{r}

library(ENMeval)
all_present <- stack(clim, efnlist)
names(all_present)
over_present <- calc.niche.overlap(all_present, "D")

names(efn245) <- names(efnlist)
efn245 <- mask(efn245, moj)
plot(efn245)

names(ant245) <- names(clim)
ant245 <- mask(ant245, moj)
plot(ant245)

names(ant370) <- names(clim)
ant370 <- mask(ant370, moj)
plot(ant370)

all_245 <- stack(ant245, efn245)
over_245 <- calc.niche.overlap(all_245, "D")

all_370 <- stack(ant370, efn370)
over_370 <- calc.niche.overlap(all_370, "D")


mean(over_245, na.omit = TRUE)
over_present <- as.data.frame(over_present)
ap_present <- dplyr::select(over_present, 1:14)
ap_present <- ap_present[15:25,]

over_245 <- as.data.frame(over_245)
ap_245 <- dplyr::select(over_245, 1:14)
ap_245 <- ap_245[15:25,]

over_370 <- as.data.frame(over_370)
ap_370 <- dplyr::select(over_370, 1:14)
ap_370 <- ap_370[15:25,]


ap_present$plant <- row.names(ap_present)
ap_present <- pivot_longer(ap_present, 1:14, names_to = "ant", values_to = "overlap")

ap_245$plant <- row.names(ap_245)
ap_245 <- pivot_longer(ap_245, 1:14, names_to = "ant", values_to = "overlap")

ap_370$plant <- row.names(ap_370)
ap_370 <- pivot_longer(ap_370, 1:14, names_to = "ant", values_to = "overlap")

mean(ap_present$overlap)
mean(ap_245$overlap)
mean(ap_370$overlap)

t.test(ap_present$overlap,ap_245$overlap, paired = TRUE)


t.test(ap_present$overlap,ap_370$overlap, paired = TRUE)

ap_present$scen <- "pres"
ap_370$scen <- "370"
ap <- cbind(ap_present, ap_370) 
colnames(ap) <- make.unique(names(ap))
ggplot(ap, aes(overlap, overlap.1)) + geom_point() + geom_abline(slope = 1) + xlim(0, 1) + ylim(0,1) + xlab("Present overlap") + ylab("Predicted overlap SSP 370")  + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"))


#370
rel <- mutate(ap, diff = (overlap.1 - overlap)/overlap)
rel <- select(rel, plant, ant, diff)

rel <- pivot_wider(rel, names_from = ant, values_from = diff)
row.names(rel) <- rel$plant
rel$plant
rel <- rel[,-1]
plantnames <- c("Cylindropuntia acanthocarpa", "Cylindropuntia echinocarpa", "Senegalia greggi", "Fouquieria splendens", "Chilopsis linearis", "Opuntia basilaris", "Cylindropuntia bigelovii", "Ferocactus cylindraceus", "Prosopis juliflora", "Prunus fascilata", "Prunus fremonti")
antnames <- c("Camponotus fragilis", "Camponotus ocreatus", "Camponotus semitestaceus", "Myrmecocystus mimicus", "Crematogaster depilis", "Dorymyrmex bicolor", "Dorymrymex insanus", "Myrmecocystus flaviceps", "Myrmecocystus kennedyi", "Myrmecocystus testaceus", "Solenopsis xyloni", "Liometopum luctuosum", "Pheidole vistana", "Forelius mccooki")
library(corrplot)
rel <- as.matrix(rel)
row.names(rel) <- plantnames
colnames(rel) <- antnames


ap245 <- cbind(ap_present, ap_245)
colnames(ap245) <- make.unique(names(ap245))
rel245 <- mutate(ap245, diff = (overlap.1 - overlap)/overlap)
rel245 <- select(rel245, plant, ant, diff)
rel245 <- pivot_wider(rel245, names_from = ant, values_from = diff)
rel245 <- rel245[,-1]
rel245 <- as.matrix(rel245)
row.names(rel245) <- plantnames
colnames(rel245) <- antnames



corrplot(rel, method = "circle", is.corr = FALSE, tl.srt = 45, tl.col = 'black', col.lim = c(-0.35, 0.35))


corrplot(rel245, method = "circle", is.corr = FALSE, tl.srt = 45, tl.col = 'black', col.lim = c(-0.35, 0.35))


```




```{r}

phydist <- cophenetic(pruned)
ant_overlap <- select(over_present, 1:14)
ant_overlap <- ant_overlap[1:14,]
names_overlap <- c("CAMPFRA", "CAMPOC", "CAMPSEM", "MYRMIM", "CREMD", "DORBI", "DORIN", "MYRFLA", "MYRKEN", "MYRTES", "SX", "LIOLUC", "PHEVIS", "FM")
colnames(ant_overlap) <- names_overlap  
row.names(ant_overlap) <- names_overlap             
phydist <- as.data.frame(phydist)
ant_overlap <- ant_overlap[names(phydist)]
ant_overlap <- 1 - ant_overlap 

ants <- ant_overlap[rownames(phydist), ]
mantel(phydist, ants, na.rm = TRUE)


ant_over370 <- select(over_370, 1:14)
ant_over370 <- ant_over370[1:14,]
colnames(ant_over370) <- names_overlap  
row.names(ant_over370) <- names_overlap

ant_over370 <- ant_over370[names(phydist)]
ant_over370 <- 1 - ant_over370
ants <- ant_over370[rownames(phydist), ]
mantel(phydist, ants, na.rm = TRUE)
```



### Levin's B Niche Breadth

```{r}
library(ENMTools)

climstack <- terra::rast(clim)





ants_breath <- data.frame(name = character(0), b1 = numeric(0), b2 = numeric(0))
for (i in 1:14){
  print(i)
  print(names(climstack[[i]]))
  plot(climstack[[i]])
  b <- raster.breadth(climstack[[i]])
  ants_breath[[i,1]] <- names(climstack[[i]])
  ants_breath[[i,2]] <- b$B1
  ants_breath[[i,3]] <- b$B2
}

efn_breath <- data.frame(name = character(0), b1 = numeric(0), b2 = numeric(0))

climstack <- terra::rast(efnlist)
for (i in 1:11){
  print(i)
  print(names(climstack[[i]]))
  plot(climstack[[i]])
  b <- raster.breadth(climstack[[i]])
  efn_breath[[i,1]] <- names(climstack[[i]])
  efn_breath[[i,2]] <- b$B1
  efn_breath[[i,3]] <- b$B2
}

check <- filter(efn_breath, name != "PFREM")
t.test(ants_breath$b2, efn_breath$b2)

t.test(ants_breath$b2, check$b2)

sd(ants_breath$b2)
sd(efn_breath$b2)
sd(check$b2)
```






## Breadth \* suitability

```{r}

ant_diff <- t(ant_diff)
ant_diff <- as.data.frame(ant_diff)
row.names(ant_diff) <- ant_diff$names
ant_diff[] <- lapply(ant_diff, as.numeric)



cor.test(ant_diff$diff245, ants_breath$b2)


cor.test(ant_diff$diff370, ants_breath$b2)


ggplot(ant_diff, aes(diff370, ants_breath$b2)) + geom_smooth(method = "lm")


efn_diff <- t(efn_diff)
efn_diff <- as.data.frame(efn_diff)
row.names(efn_diff) <- efn_diff$names
efn_diff[] <- lapply(efn_diff, as.numeric)



cor.test(efn_diff$diff245, efn_breath$b2)


cor.test(efn_diff$diff370, efn_breath$b2)


ggplot(efn_diff, aes(diff370, efn_breath$b2)) + geom_smooth(method = "lm")

efn_bre <- cbind(efn_diff, efn_breath$b2)
efn_bre$type <- "EFN"
efn_bre <- rename(efn_bre, breadth = 'efn_breath$b2')

ant_bre <- cbind(ant_diff, ants_breath$b2)
ant_bre$type <- "ANT"
ant_bre <- rename(ant_bre, breadth = 'ants_breath$b2')

all_bre <- rbind(efn_bre, ant_bre)

```

### Breadth figure

```{r}
a <- ggplot(efn_bre, aes(diff370, breadth)) + geom_smooth(method = "lm", linetype = "dashed", se = FALSE) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black")) + geom_point() + ylab("Environmental Breadth (Levin's B)") + xlab("Mean difference in suitability for EFN plants")

b <- ggplot(ant_bre, aes(diff370, breadth)) + geom_smooth(method = "lm") + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black")) + geom_point() + ylab("Environmental Breadth (Levin's B)") + xlab("Mean difference in suitability for ants")

library(cowplot)

plot_grid(a, b)
```







```{r}

traitphy <- ants_breath
traitphy$species <- c("CAMPFRA", "CAMPOC", "CAMPSEM", "MYRMIM", "CREMD", "DORBI", "DORIN", "MYRFLA", "MYRKEN", "MYRTES", "SX", "LIOLUC", "PHEVIS", "FM")




labelorder <- pruned[["tip.label"]]
labelorder

tp <- traitphy[match(labelorder, traitphy$species),]
row.names(tp) <- tp$species

traitphy <- select(tp, b2)
pruned[["tip.label"]]

#par(mfrow=c(2,2))
for (i in names(traitphy)) {
plot(pruned, show.tip.label=FALSE, main=i)
tiplabels(pch=22, col=traitphy[,i]+1, bg=traitphy[,i]+1, cex=1.5)
}

pruned <- multi2di(pruned)
multiPhylosignal(traitphy, pruned, reps = 999)

traitphy
trait <- traitphy$b2
picante::phylosignal(trait, pruned)

ps <- multiPhylosignal(traitphy, pruned)

ps



library(adephylo)

library(phylosignal)
library(phylobase)
p4d <- phylo4d(pruned, trait)
barplot.phylo4d(p4d, tree.type = "phylo", tree.ladderize = TRUE)
phyloSignal(p4d = p4d, method = "all")

ant_diff <- as.data.frame(ant_diff)
ant_diff$species <- c("CAMPFRA", "CAMPOC", "CAMPSEM", "MYRMIM", "CREMD", "DORBI", "DORIN", "MYRFLA", "MYRKEN", "MYRTES", "SX", "LIOLUC", "PHEVIS", "FM")
row.names(ant_diff) <- c("CAMPFRA", "CAMPOC", "CAMPSEM", "MYRMIM", "CREMD", "DORBI", "DORIN", "MYRFLA", "MYRKEN", "MYRTES", "SX", "LIOLUC", "PHEVIS", "FM") 


traitphy <- ant_diff



traitphy <- traitphy[match(labelorder, traitphy$species),]
traitphy <- select(traitphy, diff370)





trait <- traitphy$diff370
picante::phylosignal(trait, pruned)

traitphy <- ant_diff



traitphy <- traitphy[match(labelorder, traitphy$species),]
traitphy <- select(traitphy, diff245)
trait <- traitphy$diff245
picante::phylosignal(trait, pruned)


```


```{r}
viantlist <- list.files(path = "data/output/vi_ants/", pattern='.csv', all.files=TRUE, full.names=TRUE)

vi <- purrr::map_df(viantlist, 
              ~read.csv(.x, stringsAsFactors = FALSE) %>% mutate(filename = .x))
vi$filename <- gsub("data/output/vi_ants/", "", vi$filename)
vi <- separate(vi, filename, into = c("vi", "species", "end"), sep = "_")

vi$end <- replace_na(vi$end,"clim")
vi <- filter(vi, end == "clim")

#remove the two excluded ant species (due to poor AUC performance)
vi <- filter(vi, species != "fp.csv" & species != "campsay.csv")

ggplot(vi, aes(Variable, Permutation_importance)) + geom_boxplot()


pltvilist <-  list.files(path = "data/output/vi/", pattern='.csv', all.files=TRUE, full.names=TRUE)
pltvi <- purrr::map_df(pltvilist, 
              ~read.csv(.x, stringsAsFactors = FALSE) %>% mutate(filename = .x))
pltvi$filename <-  gsub("data/output/vi/", "", pltvi$filename)
pltvi$filename <-  gsub(".csv", "", pltvi$filename)
pltvi$filename <-  gsub("vi_", "", pltvi$filename)

pltvi <- filter(pltvi, filename == "ag" |filename == "ca" | filename == "ce" | filename == "cl" | filename == "fa" | filename == "ob" | filename == "obig" | filename == "pfrem" | filename == "fa" | filename == "fs" | filename == "pj")


pltvi$type <- "EFN"
vi$type <- "ant"
pltvi <- select(pltvi, -sd)
vi <- select(vi, 1:3, 6, 8)
pltvi <- rename(pltvi, species = filename)
bothvi <- rbind(pltvi, vi)

ggplot(bothvi, aes(Variable, Permutation_importance, fill = type)) + geom_boxplot()

pc1 <- filter(bothvi, Variable == "PC1")
t.test(Permutation_importance ~ type, data = pc1)
pc2 <- filter(bothvi, Variable == "PC2")
t.test(Permutation_importance ~ type, data = pc2)
pc3 <- filter(bothvi, Variable == "PC3")
t.test(Permutation_importance ~ type, data = pc3)
pc4 <- filter(bothvi, Variable == "PC4")
t.test(Permutation_importance ~ type, data = pc4)

```


